---
title: "House Prices Lasso Regression"
author: "Autumn Brinkerhoff"
date: '2024-01-05'
output: pdf_document
---

##Loading libraries
```{r }

library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
##library(xgboost)

```

##Reading the data
```{r }

train <- read.csv("train.csv")
test <- read.csv("test.csv")

dim(train)
dim(test)

```


##Data structure
```{r }

test_labels <- test$Id
test$Id <- NULL
train$Id <- NULL

test$SalePrice <- NA
df <- rbind(train, test)
dim(df)

head(df)
summary(df)
```

```{r}

ggplot(data=df[!is.na(df$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="blue", binwidth = 10000) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000), labels = comma)

summary(df$SalePrice)
```

```{r}

numericVars <- which(sapply(df, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
cat('There are', length(numericVars), 'numeric variables')

```



```{r}
df_numVar <- df[, numericVars]
cor_numVar <- cor(df_numVar, use="pairwise.complete.obs") #correlations of df numeric variables with NAs


cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))

CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.4)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")
```


```{r}
ggplot(data=df[!is.na(df$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot() + labs(x='Overdf Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```


```{r}
ggplot(data=df[!is.na(df$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_text_repel(aes(label = ifelse(df$GrLivArea[!is.na(df$SalePrice)]>4500, rownames(df), '')))
```


```{r}

df[c(524, 1299), c('SalePrice', 'GrLivArea', 'OverallQual')]

NAcol <- which(colSums(is.na(df)) > 0)
sort(colSums(sapply(df[NAcol], is.na)), decreasing = TRUE)

```
##missing data

```{r}

df$PoolQC[is.na(df$PoolQC)] <- 'None'

Qualities <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)

df$PoolQC<-as.integer(revalue(df$PoolQC, Qualities))
table(df$PoolQC)

df[df$PoolArea>0 & df$PoolQC==0, c('PoolArea', 'PoolQC', 'OverallQual')]

```

```{r}
df$PoolQC[2421] <- 2
df$PoolQC[2504] <- 3
df$PoolQC[2600] <- 2


ggplot(df[!is.na(df$SalePrice),], aes(x=PoolArea, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..))
```

##Miscellaneous Feature
```{r}

df$MiscFeature[is.na(df$MiscFeature)] <- 'None'
df$MiscFeature <- as.factor(df$MiscFeature)

ggplot(df[!is.na(df$SalePrice),], aes(x=MiscFeature, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..))

table(df$MiscFeature)

```
##Alley
```{r}

df$Alley[is.na(df$Alley)] <- 'None'
df$Alley <- as.factor(df$Alley)

ggplot(df[!is.na(df$SalePrice),], aes(x=Alley, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue')+
        scale_y_continuous(breaks= seq(0, 200000, by=50000), labels = comma)

table(df$Alley)
```
##FireplaceQu
```{r}
df$FireplaceQu[is.na(df$FireplaceQu)] <- 'None'
df$FireplaceQu<-as.integer(revalue(df$FireplaceQu, Qualities))
table(df$FireplaceQu)
```


```{r}
ggplot(df[!is.na(df$LotFrontage),], aes(x=as.factor(Neighborhood), y=LotFrontage)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

for (i in 1:nrow(df)){
        if(is.na(df$LotFrontage[i])){
               df$LotFrontage[i] <- as.integer(median(df$LotFrontage[df$Neighborhood==df$Neighborhood[i]], na.rm=TRUE)) 
        }
}


df$LotShape<-as.integer(revalue(df$LotShape, c('IR3'=0, 'IR2'=1, 'IR1'=2, 'Reg'=3)))
table(df$LotShape)
```

```{r}
length(which(is.na(df$GarageType) & is.na(df$GarageFinish) & is.na(df$GarageCond) & is.na(df$GarageQual)))

kable(df[!is.na(df$GarageType) & is.na(df$GarageFinish), c('GarageCars', 'GarageArea', 'GarageType', 'GarageCond', 'GarageQual', 'GarageFinish')])

df$GarageCond[2127] <- names(sort(-table(df$GarageCond)))[1]
df$GarageQual[2127] <- names(sort(-table(df$GarageQual)))[1]
df$GarageFinish[2127] <- names(sort(-table(df$GarageFinish)))[1]

#display "fixed" house
kable(df[2127, c('GarageYrBlt', 'GarageCars', 'GarageArea', 'GarageType', 'GarageCond', 'GarageQual', 'GarageFinish')])
```

```{r}
#fixing 3 values for house 2577
df$GarageCars[2577] <- 0
df$GarageArea[2577] <- 0
df$GarageType[2577] <- NA

#check if NAs of the character variables are now df 158
length(which(is.na(df$GarageType) & is.na(df$GarageFinish) & is.na(df$GarageCond) & is.na(df$GarageQual)))
```

```{r}
df$GarageType[is.na(df$GarageType)] <- 'No Garage'
df$GarageType <- as.factor(df$GarageType)
table(df$GarageType)

df$GarageFinish[is.na(df$GarageFinish)] <- 'None'
Finish <- c('None'=0, 'Unf'=1, 'RFn'=2, 'Fin'=3)

df$GarageFinish<-as.integer(revalue(df$GarageFinish, Finish))
table(df$GarageFinish)
```
```{r}
df$GarageQual[is.na(df$GarageQual)] <- 'None'
df$GarageQual<-as.integer(revalue(df$GarageQual, Qualities))
table(df$GarageQual)

df$GarageCond[is.na(df$GarageCond)] <- 'None'
df$GarageCond<-as.integer(revalue(df$GarageCond, Qualities))
table(df$GarageCond)
```

##Basement
```{r}

length(which(is.na(df$BsmtQual) & is.na(df$BsmtCond) & is.na(df$BsmtExposure) & is.na(df$BsmtFinType1) & is.na(df$BsmtFinType2)))

```

```{r}
#Find the additional NAs; BsmtFinType1 is the one with 79 NAs
df[!is.na(df$BsmtFinType1) & (is.na(df$BsmtCond)|is.na(df$BsmtQual)|is.na(df$BsmtExposure)|is.na(df$BsmtFinType2)), c('BsmtQual', 'BsmtCond', 'BsmtExposure', 'BsmtFinType1', 'BsmtFinType2')]

#Imputing modes.
df$BsmtFinType2[333] <- names(sort(-table(df$BsmtFinType2)))[1]
df$BsmtExposure[c(949, 1488, 2349)] <- names(sort(-table(df$BsmtExposure)))[1]
df$BsmtCond[c(2041, 2186, 2525)] <- names(sort(-table(df$BsmtCond)))[1]
df$BsmtQual[c(2218, 2219)] <- names(sort(-table(df$BsmtQual)))[1]
```

```{r}
df$BsmtQual[is.na(df$BsmtQual)] <- 'None'
df$BsmtQual<-as.integer(revalue(df$BsmtQual, Qualities))
table(df$BsmtQual)
```


```{r}
df$BsmtCond[is.na(df$BsmtCond)] <- 'None'
df$BsmtCond<-as.integer(revalue(df$BsmtCond, Qualities))
table(df$BsmtCond)

df$BsmtExposure[is.na(df$BsmtExposure)] <- 'None'
Exposure <- c('None'=0, 'No'=1, 'Mn'=2, 'Av'=3, 'Gd'=4)

df$BsmtExposure<-as.integer(revalue(df$BsmtExposure, Exposure))
table(df$BsmtExposure)

df$BsmtFinType1[is.na(df$BsmtFinType1)] <- 'None'
FinType <- c('None'=0, 'Unf'=1, 'LwQ'=2, 'Rec'=3, 'BLQ'=4, 'ALQ'=5, 'GLQ'=6)

df$BsmtFinType1<-as.integer(revalue(df$BsmtFinType1, FinType))
table(df$BsmtFinType1)

df$BsmtFinType2[is.na(df$BsmtFinType2)] <- 'None'
FinType <- c('None'=0, 'Unf'=1, 'LwQ'=2, 'Rec'=3, 'BLQ'=4, 'ALQ'=5, 'GLQ'=6)

df$BsmtFinType2<-as.integer(revalue(df$BsmtFinType2, FinType))
table(df$BsmtFinType2)
```


```{r}
#display remaining NAs. Using BsmtQual as a reference for the 79 houses without basement agreed upon earlier
df[(is.na(df$BsmtFullBath)|is.na(df$BsmtHalfBath)|is.na(df$BsmtFinSF1)|is.na(df$BsmtFinSF2)|is.na(df$BsmtUnfSF)|is.na(df$TotalBsmtSF)), c('BsmtQual', 'BsmtFullBath', 'BsmtHalfBath', 'BsmtFinSF1', 'BsmtFinSF2', 'BsmtUnfSF', 'TotalBsmtSF')]

df$BsmtFullBath[is.na(df$BsmtFullBath)] <-0
table(df$BsmtFullBath)

df$BsmtHalfBath[is.na(df$BsmtHalfBath)] <-0
table(df$BsmtHalfBath)

df$BsmtFinSF1[is.na(df$BsmtFinSF1)] <-0

df$BsmtFinSF2[is.na(df$BsmtFinSF2)] <-0

df$BsmtUnfSF[is.na(df$BsmtUnfSF)] <-0

df$TotalBsmtSF[is.na(df$TotalBsmtSF)] <-0
```
##Masonry
```{r}
#check if the 23 houses with veneer area NA are also NA in the veneer type
length(which(is.na(df$MasVnrType) & is.na(df$MasVnrArea)))

df[is.na(df$MasVnrType) & !is.na(df$MasVnrArea), c('MasVnrType', 'MasVnrArea')]

#fix this veneer type by imputing the mode
df$MasVnrType[2611] <- names(sort(-table(df$MasVnrType)))[2] #taking the 2nd value as the 1st is 'none'
df[2611, c('MasVnrType', 'MasVnrArea')]

df$MasVnrType[is.na(df$MasVnrType)] <- 'None'

df[!is.na(df$SalePrice),] %>% group_by(MasVnrType) %>% summarise(median = median(SalePrice), counts=n()) %>% arrange(median)

Masonry <- c('None'=0, 'BrkCmn'=0, 'BrkFace'=1, 'Stone'=2)
df$MasVnrType<-as.integer(revalue(df$MasVnrType, Masonry))
table(df$MasVnrType)

df$MasVnrArea[is.na(df$MasVnrArea)] <-0
```

##MSZoning
```{r}
#imputing the mode
df$MSZoning[is.na(df$MSZoning)] <- names(sort(-table(df$MSZoning)))[1]
df$MSZoning <- as.factor(df$MSZoning)
table(df$MSZoning)
```

##Kitchen
```{r}
df$KitchenQual[is.na(df$KitchenQual)] <- 'TA' #replace with most common value
df$KitchenQual<-as.integer(revalue(df$KitchenQual, Qualities))
table(df$KitchenQual)

table(df$KitchenAbvGr)
```
##Utilities
```{r}
table(df$Utilities)

kable(df[is.na(df$Utilities) | df$Utilities=='NoSeWa', 1:9])

df$Utilities <- NULL
```

##Home functionality
```{r}
#impute mode for the 1 NA
df$Functional[is.na(df$Functional)] <- names(sort(-table(df$Functional)))[1]

df$Functional <- as.integer(revalue(df$Functional, c('Sal'=0, 'Sev'=1, 'Maj2'=2, 'Maj1'=3, 'Mod'=4, 'Min2'=5, 'Min1'=6, 'Typ'=7)))
table(df$Functional)
```

##exterior variables
```{r}
#imputing mode
df$Exterior1st[is.na(df$Exterior1st)] <- names(sort(-table(df$Exterior1st)))[1]

df$Exterior1st <- as.factor(df$Exterior1st)
table(df$Exterior1st)


#imputing mode
df$Exterior2nd[is.na(df$Exterior2nd)] <- names(sort(-table(df$Exterior2nd)))[1]

df$Exterior2nd <- as.factor(df$Exterior2nd)
table(df$Exterior2nd)


df$ExterCond<-as.integer(revalue(df$ExterCond, Qualities))

sum(table(df$ExterCond))
```

##Electrical system
```{r}
#imputing mode
df$Electrical[is.na(df$Electrical)] <- names(sort(-table(df$Electrical)))[1]

df$Electrical <- as.factor(df$Electrical)
table(df$Electrical)

sum(table(df$Electrical))
```

##Fence
```{r}
df$Fence[is.na(df$Fence)] <- 'None'
table(df$Fence)

df[!is.na(df$SalePrice),] %>% group_by(Fence) %>% summarise(median = median(SalePrice), counts=n())

df$Fence <- as.factor(df$Fence)
```


##SaleType
```{r}
#imputing mode
df$SaleType[is.na(df$SaleType)] <- names(sort(-table(df$SaleType)))[1]

df$SaleType <- as.factor(df$SaleType)
table(df$SaleType)

df$SaleCondition <- as.factor(df$SaleCondition)
table(df$SaleCondition)

sum(table(df$SaleCondition))
```



```{r}
Charcol <- names(df[,sapply(df, is.character)])
Charcol

cat('There are', length(Charcol), 'remaining columns with character values')
```
##Foundation
```{r}
#No ordinality, so converting into factors
df$Foundation <- as.factor(df$Foundation)
table(df$Foundation)

sum(table(df$Foundation))
```

##Heating
```{r}
#No ordinality, so converting into factors
df$Heating <- as.factor(df$Heating)
table(df$Heating)

sum(table(df$Heating))
```

##RoofStyle
```{r}
#No ordinality, so converting into factors
df$RoofStyle <- as.factor(df$RoofStyle)
table(df$RoofStyle)

sum(table(df$RoofStyle))
```


##LandContour
```{r}
#No ordinality, so converting into factors
df$LandContour <- as.factor(df$LandContour)
table(df$LandContour)

sum(table(df$LandContour))
```


##BldgType
```{r}
#No ordinality, so converting into factors
df$BldgType <- as.factor(df$BldgType)
table(df$BldgType)

sum(table(df$BldgType))
```


##Neighborhood
```{r}
#No ordinality, so converting into factors
df$Neighborhood <- as.factor(df$Neighborhood)
table(df$Neighborhood)

sum(table(df$Neighborhood))
```

##Street
```{r}
#Ordinal, so label encoding
df$Street<-as.integer(revalue(df$Street, c('Grvl'=0, 'Pave'=1)))
table(df$Street)

sum(table(df$Street))
```

```{r}
#Ordinal, so label encoding
df$PavedDrive<-as.integer(revalue(df$PavedDrive, c('N'=0, 'P'=1, 'Y'=2)))
table(df$PavedDrive)

sum(table(df$PavedDrive))
```

```{r}
df$MoSold <- as.factor(df$MoSold)
```

```{r}
ys <- ggplot(df[!is.na(df$SalePrice),], aes(x=as.factor(YrSold), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue')+
        scale_y_continuous(breaks= seq(0, 800000, by=25000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..)) +
        coord_cartesian(ylim = c(0, 200000)) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice

ms <- ggplot(df[!is.na(df$SalePrice),], aes(x=MoSold, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue')+
        scale_y_continuous(breaks= seq(0, 800000, by=25000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..)) +
        coord_cartesian(ylim = c(0, 200000)) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice

grid.arrange(ys, ms, widths=c(1,2))
```


```{r}
df$MSSubClass <- as.factor(df$MSSubClass)

#revalue for better readability
df$MSSubClass<-revalue(df$MSSubClass, c('20'='1 story 1946+', '30'='1 story 1945-', '40'='1 story unf attic', '45'='1,5 story unf', '50'='1,5 story fin', '60'='2 story 1946+', '70'='2 story 1945-', '75'='2,5 story df ages', '80'='split/multi level', '85'='split foyer', '90'='duplex df style/age', '120'='1 story PUD 1946+', '150'='1,5 story PUD df', '160'='2 story PUD 1946+', '180'='PUD multilevel', '190'='2 family conversion'))

str(df$MSSubClass)
```


#Visualization
```{r}
numericVars <- which(sapply(df, is.numeric)) #index vector numeric variables
factorVars <- which(sapply(df, is.factor)) #index vector factor variables
cat('There are', length(numericVars), 'numeric variables, and', length(factorVars), 'categoric variables')
```


#Correlations
```{r}
df_numVar <- df[, numericVars]
cor_numVar <- cor(df_numVar, use="pairwise.complete.obs") #correlations of df numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt", tl.cex = 0.7,cl.cex = .7, number.cex=.7)
```

```{r}
#is.na(df$SalePrice)
```

```{r}
s1 <- ggplot(data= df, aes(x=GrLivArea)) +
        geom_density() + labs(x='Square feet living area')
s2 <- ggplot(data=df, aes(x=as.factor(TotRmsAbvGrd))) +
        geom_histogram(stat='count') + labs(x='Rooms above Ground')
s3 <- ggplot(data= df, aes(x=X1stFlrSF)) +
        geom_density() + labs(x='Square feet first floor')
s4 <- ggplot(data= df, aes(x=X2ndFlrSF)) +
        geom_density() + labs(x='Square feet second floor')
s5 <- ggplot(data= df, aes(x=TotalBsmtSF)) +
        geom_density() + labs(x='Square feet basement')
s6 <- ggplot(data= df[df$LotArea<100000,], aes(x=LotArea)) +
        geom_density() + labs(x='Square feet lot')
s7 <- ggplot(data= df, aes(x=LotFrontage)) +
        geom_density() + labs(x='Linear feet lot frontage')
s8 <- ggplot(data= df, aes(x=LowQualFinSF)) +
        geom_histogram() + labs(x='Low quality square feet 1st & 2nd')

layout <- matrix(c(1,2,5,3,4,8,6,7),4,2,byrow=TRUE)
multiplot(s1, s2, s3, s4, s5, s6, s7, s8, layout=layout)
```


```{r}
cor(df$GrLivArea, (df$X1stFlrSF + df$X2ndFlrSF + df$LowQualFinSF))
```


```{r}
head(df[df$LowQualFinSF>0, c('GrLivArea', 'X1stFlrSF', 'X2ndFlrSF', 'LowQualFinSF')])

n1 <- ggplot(df[!is.na(df$SalePrice),], aes(x=Neighborhood, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
n2 <- ggplot(data=df, aes(x=Neighborhood)) +
        geom_histogram(stat='count')+
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3)+
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(n1, n2)
```


```{r}
q1 <- ggplot(data=df, aes(x=as.factor(OverallQual))) +
        geom_histogram(stat='count')
q2 <- ggplot(data=df, aes(x=as.factor(ExterQual))) +
        geom_histogram(stat='count')
q3 <- ggplot(data=df, aes(x=as.factor(BsmtQual))) +
        geom_histogram(stat='count')
q4 <- ggplot(data=df, aes(x=as.factor(KitchenQual))) +
        geom_histogram(stat='count')
q5 <- ggplot(data=df, aes(x=as.factor(GarageQual))) +
        geom_histogram(stat='count')
q6 <- ggplot(data=df, aes(x=as.factor(FireplaceQu))) +
        geom_histogram(stat='count')
q7 <- ggplot(data=df, aes(x=as.factor(PoolQC))) +
        geom_histogram(stat='count')

layout <- matrix(c(1,2,8,3,4,8,5,6,7),3,3,byrow=TRUE)
multiplot(q1, q2, q3, q4, q5, q6, q7, layout=layout)
```


```{r}
ms1 <- ggplot(df[!is.na(df$SalePrice),], aes(x=MSSubClass, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
ms2 <- ggplot(data=df, aes(x=MSSubClass)) +
        geom_histogram(stat='count')+
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(ms1, ms2)
```


```{r}
#correct error
df$GarageYrBlt[2593] <- 2007 #this must have been a typo. GarageYrBlt=2207, YearBuilt=2006, YearRemodAdd=2007.

g1 <- ggplot(data=df[df$GarageCars !=0,], aes(x=GarageYrBlt)) +
        geom_histogram()
g2 <- ggplot(data=df, aes(x=as.factor(GarageCars))) +
        geom_histogram(stat='count')
g3 <- ggplot(data= df, aes(x=GarageArea)) +
        geom_density()
g4 <- ggplot(data=df, aes(x=as.factor(GarageCond))) +
        geom_histogram(stat='count')
g5 <- ggplot(data=df, aes(x=GarageType)) +
        geom_histogram(stat='count')
g6 <- ggplot(data=df, aes(x=as.factor(GarageQual))) +
        geom_histogram(stat='count')
g7 <- ggplot(data=df, aes(x=as.factor(GarageFinish))) +
        geom_histogram(stat='count')

layout <- matrix(c(1,5,5,2,3,8,6,4,7),3,3,byrow=TRUE)
multiplot(g1, g2, g3, g4, g5, g6, g7, layout=layout)
```



```{r}
b1 <- ggplot(data=df, aes(x=BsmtFinSF1)) +
        geom_histogram() + labs(x='Type 1 finished square feet')
b2 <- ggplot(data=df, aes(x=BsmtFinSF2)) +
        geom_histogram()+ labs(x='Type 2 finished square feet')
b3 <- ggplot(data=df, aes(x=BsmtUnfSF)) +
        geom_histogram()+ labs(x='Unfinished square feet')
b4 <- ggplot(data=df, aes(x=as.factor(BsmtFinType1))) +
        geom_histogram(stat='count')+ labs(x='Rating of Type 1 finished area')
b5 <- ggplot(data=df, aes(x=as.factor(BsmtFinType2))) +
        geom_histogram(stat='count')+ labs(x='Rating of Type 2 finished area')
b6 <- ggplot(data=df, aes(x=as.factor(BsmtQual))) +
        geom_histogram(stat='count')+ labs(x='Height of the basement')
b7 <- ggplot(data=df, aes(x=as.factor(BsmtCond))) +
        geom_histogram(stat='count')+ labs(x='Rating of general condition')
b8 <- ggplot(data=df, aes(x=as.factor(BsmtExposure))) +
        geom_histogram(stat='count')+ labs(x='Walkout or garden level wdfs')

layout <- matrix(c(1,2,3,4,5,9,6,7,8),3,3,byrow=TRUE)
multiplot(b1, b2, b3, b4, b5, b6, b7, b8, layout=layout)

```



```{r}
df$TotBathrooms <- df$FullBath + (df$HalfBath*0.5) + df$BsmtFullBath + (df$BsmtHalfBath*0.5)

tb1 <- ggplot(data=df[!is.na(df$SalePrice),], aes(x=as.factor(TotBathrooms), y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
tb2 <- ggplot(data=df, aes(x=as.factor(TotBathrooms))) +
        geom_histogram(stat='count')
grid.arrange(tb1, tb2)
```



```{r}
df$Remod <- ifelse(df$YearBuilt==df$YearRemodAdd, 0, 1) #0=No Remodeling, 1=Remodeling
df$Age <- as.numeric(df$YrSold)-df$YearRemodAdd

ggplot(data=df[!is.na(df$SalePrice),], aes(x=Age, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```



```{r}
cor(df$SalePrice[!is.na(df$SalePrice)], df$Age[!is.na(df$SalePrice)])

ggplot(df[!is.na(df$SalePrice),], aes(x=as.factor(Remod), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=6) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        theme_grey(base_size = 18) +
        geom_hline(yintercept=163000, linetype="dashed") #dashed line is median SalePrice
```


```{r}
df$IsNew <- ifelse(df$YrSold==df$YearBuilt, 1, 0)
table(df$IsNew)

```


```{r}
ggplot(df[!is.na(df$SalePrice),], aes(x=as.factor(IsNew), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=6) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        theme_grey(base_size = 18) +
        geom_hline(yintercept=163000, linetype="dashed") #dashed line is median SalePrice
```


```{r}
df$YrSold <- as.factor(df$YrSold) #the numeric version is now not needed anymore
```

```{r}
nb1 <- ggplot(df[!is.na(df$SalePrice),], aes(x=reorder(Neighborhood, SalePrice, FUN=median), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') + labs(x='Neighborhood', y='Median SalePrice') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
nb2 <- ggplot(df[!is.na(df$SalePrice),], aes(x=reorder(Neighborhood, SalePrice, FUN=mean), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "mean", fill='blue') + labs(x='Neighborhood', y="Mean SalePrice") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
grid.arrange(nb1, nb2)
```


```{r}
df$NeighRich[df$Neighborhood %in% c('StoneBr', 'NridgHt', 'NoRidge')] <- 2
df$NeighRich[!df$Neighborhood %in% c('MeadowV', 'IDOTRR', 'BrDale', 'StoneBr', 'NridgHt', 'NoRidge')] <- 1
df$NeighRich[df$Neighborhood %in% c('MeadowV', 'IDOTRR', 'BrDale')] <- 0

table(df$NeighRich)
```



```{r}
df$TotalSqFeet <- df$GrLivArea + df$TotalBsmtSF

ggplot(data=df[!is.na(df$SalePrice),], aes(x=TotalSqFeet, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_text_repel(aes(label = ifelse(df$GrLivArea[!is.na(df$SalePrice)]>4500, rownames(df), '')))
```

```{r}
cor(df$SalePrice, df$TotalSqFeet, use= "pairwise.complete.obs")
```


```{r}
cor(df$SalePrice[-c(524, 1299)], df$TotalSqFeet[-c(524, 1299)], use= "pairwise.complete.obs")


```


```{r}
df$TotalPorchSF <- df$OpenPorchSF + df$EnclosedPorch + df$X3SsnPorch + df$ScreenPorch

cor(df$SalePrice, df$TotalPorchSF, use= "pairwise.complete.obs")

ggplot(data=df[!is.na(df$SalePrice),], aes(x=TotalPorchSF, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

#Preparing data for modeling
```{r}

dropVars <- c('YearRemodAdd', 'GarageYrBlt', 'GarageArea', 'GarageCond', 'TotalBsmtSF', 'TotalRmsAbvGrd', 'BsmtFinSF1')

df <- df[,!(names(df) %in% dropVars)]


df <- df[-c(524, 1299),]
```

```{r}
numericVarNames <- numericVarNames[!(numericVarNames %in% c('MSSubClass', 'MoSold', 'YrSold', 'SalePrice', 'OverdfQual', 'OverdfCond'))] #numericVarNames was created before having done anything
numericVarNames <- append(numericVarNames, c('Age', 'TotalPorchSF', 'TotBathrooms', 'TotalSqFeet'))

DFnumeric <- df[, names(df) %in% numericVarNames]

DFfactors <- df[, !(names(df) %in% numericVarNames)]
DFfactors <- DFfactors[, names(DFfactors) != 'SalePrice']

cat('There are', length(DFnumeric), 'numeric variables, and', length(DFfactors), 'factor variables')
```



```{r}
for(i in 1:ncol(DFnumeric)){
        if (abs(skew(DFnumeric[,i]))>0.8){
                DFnumeric[,i] <- log(DFnumeric[,i] +1)
        }
}


PreNum <- preProcess(DFnumeric, method=c("center", "scale"))
print(PreNum)
```


```{r}
DFnorm <- predict(PreNum, DFnumeric)
dim(DFnorm)
```


```{r}
DFdummies <- as.data.frame(model.matrix(~.-1, DFfactors))
dim(DFdummies)


#check if some values are absent in the test set
ZerocolTest <- which(colSums(DFdummies[(nrow(df[!is.na(df$SalePrice),])+1):nrow(df),])==0)
colnames(DFdummies[ZerocolTest])

DFdummies <- DFdummies[,-ZerocolTest] #removing predictors
```


```{r}
#check if some values are absent in the train set
ZerocolTrain <- which(colSums(DFdummies[1:nrow(df[!is.na(df$SalePrice),]),])==0)
colnames(DFdummies[ZerocolTrain])

DFdummies <- DFdummies[,-ZerocolTrain] #removing predictor

fewOnes <- which(colSums(DFdummies[1:nrow(df[!is.na(df$SalePrice),]),])<10)
colnames(DFdummies[fewOnes])
```



```{r}
DFdummies <- DFdummies[,-fewOnes] #removing predictors
dim(DFdummies)

combined <- cbind(DFnorm, DFdummies) #combining df (now numeric) predictors into one dataframe 
```

##Dealing with skewness of response variable
```{r}
skew(df$SalePrice)

qqnorm(df$SalePrice)
qqline(df$SalePrice)
```


```{r}
df$SalePrice <- log(df$SalePrice) #default is the natural logarithm, "+1" is not necessary as there are no 0's
skew(df$SalePrice)

qqnorm(df$SalePrice)
qqline(df$SalePrice)
```

```{r}
train1 <- combined[!is.na(df$SalePrice),]
test1 <- combined[is.na(df$SalePrice),]

```


#Lasso regression model
```{r}
set.seed(27042018)

my_control <-trainControl(method="cv", number=5)
lassoGrid <- expand.grid(alpha = 1, lambda = seq(0.001,0.1,by = 0.0005))

lasso_mod <- train(x=train1, y= df$SalePrice[!is.na(df$SalePrice)], 
                   method='glmnet', 
                   trControl= my_control,
                   tuneGrid=lassoGrid) 
lasso_mod$bestTune



```

```{r}

min(lasso_mod$results$RMSE)

#print(lasso_mod$results)
summary(lasso_mod$results)

```


```{r}
lassoVarImp <- varImp(lasso_mod,scale=F)
lassoImportance <- lassoVarImp$importance

varsSelected <- length(which(lassoImportance$Overdf!=0))
varsNotSelected <- length(which(lassoImportance$Overdf==0))

cat('Lasso uses', varsSelected, 'variables in its model, and did not select', varsNotSelected, 'variables.')
```



```{r}
LassoPred <- predict(lasso_mod, test1)
predictions_lasso <- exp(LassoPred) #need to reverse the log to the real values
head(predictions_lasso)


```



